Rewind: Automatically Generated\\Digital Memories from Mobile Geolocation Data

### 2016-09-12 paper/uist_2016/paper.tex
memorability findings

Do you remember these locations? Do you remember this trip? Can  you name some of the streets? Can you name what is around these locations, such as buildings, landmarks, or other places?

@@

\section{Findings}
The findings of our user study center on the participants' relationship to the Rewinds generated from their location data. We focus on the participants' judgments of their Rewinds' memorability, believability, and desirability in comparison to their judgments of the photos paired with the Rewinds. We also discuss the participants' interactions with and reactions to their Rewinds. The results use data from the user studies conducted with 12 men and 6 women. Half of the participants identified themselves as frequent travelers. Additionally, half of the participants were regular GPS users, or people who used GPS even in locations already familiar to them.

\subsection{Memorability}
Rewinds are meant to engage a person's memory. To see if the Rewind system selected and portrayed memorable locations and trips, we investigated whether participants were able to remember the location shown in a photo or recall the trip portrayed by a Rewind. Location memorability is determined by whether or not the participant remembers the location depicted in a photo. Trip memorability is determined by whether or not the participant remembers and can distinguish the exact trip portrayed in a Rewind. In this section we will look at differences in location and trip remembrance between participants who are regular GPS users and participants who do not use GPS regularly, and examine the effect of location distance and visit frequency on memorability.

\subsubsection{Rewinds and Photos}
When participants were shown a street side photo from their location history, they were asked whether they recognized the location in the photo. In 50\% of cases, participants said they did. When they were shown the Rewind associated with that photo, nearly all participants knew the location captured---97\%. This result shows the impact of Rewind on memorability, as the addition images before and after the static photograph greatly impacts whether a past location is familiar to a user.

@@

\subsubsection{Distance}
The distance away from the participant’s home had differing effects on location memorability and trip memorability (Figure~\ref{fig:distancememory}). Distances ranged from 0 to 206,155 meters. For analysis, these distances were logarithmically scaled and normalized into six distance groups to mimic human categorization of distances into generalizations like ``close,'' ``far,'' or ``very far.'' When normalized: Group 0 corresponded to a range of 0--10 meters from user's home location; Group 1 corresponded to a range of 10--100 meters; Group 2 corresponded to a range of 100--1,000 meters; Group 3 corresponded to a range of 1,000--10,000 meters; Group 4 corresponded to a range of 10,000--100,000 meters and Group 5 corresponded to a range of 100,000--1,000,000 meters from user's home location. As the distance increased, location memorability decreased, suggesting that locations further away from home were less likely to be remembered. Location remembrance dropped from 100\% for Distance Group 0, or home, to 74\% for Distance Group 1 down to 10\% for Distance Group 5, which represented the locations furthest away from home. For trips, remembrance was largely the same for most distance groups, except for distances that were very close to home and distances that were very far, but not the furthest, from home. Trip remembrance was at 60\% for Distance Group 0, then dipped momentarily to 25\% in Distance Group 1 before increasing back up to 60\% for Distance Group 2, 3, and 5. Only Distance Group 4 wavered with a trip remembrance rate of 48\%. This data suggests that participants often did not remember or could not distinguish the exact trips occurring in locations very close to home, where most trips happen, nor could they distinguish the exact trips that happened very far, but not especially far, from home.

>>>>

Do you remember these locations? Do you remember this particular trip? Can  you name some of the streets? Can you name what is around these locations, such as buildings, landmarks, or other places?

@@

\section{Findings}
The findings of our user study center on the participants' relationship to the Rewinds generated from their location data. We focus on the participants' judgments of their Rewinds' memorability, believability, and desirability in comparison to their judgments of the photos paired with the Rewinds. We also discuss the participants' interactions with and reactions to their Rewinds. The results use data from the user studies conducted with 12 men and 6 women. Half of the participants identified themselves as frequent travelers. Additionally, half of the participants were regular GPS users, that is people who used GPS even in locations already familiar to them.

\subsection{Memorability}
Rewinds are meant to engage a person's memory. To find out more about what features make a digital memory memorable, we examined participant responses to whether or not they were able to remember the location shown in a photo or recall the trip portrayed by a Rewind. Photo memorability is determined by whether or not the participant remembered the location depicted in a photo. Because a Rewind contains many locations, it would be unfair to consider the memorability of Rewind based on whether participant remembers any locations in the Rewind. So, Rewind memorability is determined by whether or not the participant remembered and could distinguish the particular trip portrayed in a Rewind. In this section, we will look at differences in photo and Rewind memorability based on the variables of distance from the participant's home, visit frequency, presence of travel companions, regular GPS usage, and whether the participant considered themselves a frequent traveler.

\subsubsection{Rewinds versus Photos}
When participants were shown a street side photo from their location history, they were asked whether they recognized the location in the photo. In 50\% of cases, participants said they did. When they were shown the Rewind associated with that photo, nearly every case had recognizable locations---97\%. This result is expected since Rewinds contain more location information than static photos, so if participants do not recognize some depicted locations, they still have many other chances to pick out locations they remember. Rewind memorability as calculated by whether participants could identify the particular trip they were on from watching the Rewind was 54\%. However, there were cases where the participant would only remember the photo after seeing its accompanying Rewind, which provided a context. More than one participant exclaimed after watching a Rewind, ``I recognize the photo now!''

\subsubsection{Distance}
The distance away from the participant’s home had differing effects on photo memorability and Rewind memorability (Figure~\ref{fig:distancememory}). Distances ranged from 0 to 206,155 meters. For analysis, these distances were logarithmically scaled and normalized into six distance groups to mimic human categorization of distances into generalizations like ``close,'' ``far,'' or ``very far.'' When normalized: Group 0 corresponded to a range of 0--10 meters from the participant's home location; Group 1 corresponded to a range of 10--100 meters; Group 2 corresponded to a range of 100--1,000 meters; Group 3 corresponded to a range of 1,000--10,000 meters; Group 4 corresponded to a range of 10,000--100,000 meters and Group 5 corresponded to a range of 100,000--1,000,000 meters from the participant's home location.

Photo memorability and distance had a clear linear relationship. As the distance increased, photo memorability decreased, suggesting that locations further away from home were less likely to be remembered. Remembrance was 100\% for Distance Group 0, or home. One participant recognized a photo because her ``husband's car is in the driveway.'' Remembrance dropped to 74\% for Distance Group 1, then down to 10\% for Distance Group 5, which represented the locations furthest away from home. These percentages are statistically significant, with a p-value of 0.0028.

There was no statistical significance for Rewind remembrance. Our data does show, however, that Rewind remembrance was largely the same for most distance groups, except for distances that were close to home. While remembrance for almost all distance groups was on average near 60\%, remembrance for Rewinds in Distance Group 1 was 25\%. This data suggests that participants often did not remember or could not distinguish the exact trips occurring in locations close to their home, where most trips happen.

@@

\subsubsection{Frequent Travelers}
Our findings show that trip remembrance is influenced by the presence of travel companions. As one might imagine, trips that were taken with other people were much more memorable than trips taken alone. Participants were not able to remember the exact trip of 33\% of Rewinds taken when they were traveling alone, but they could remember the exact trip of 89\% of Rewinds that were taken with company.

\subsubsection{Other Variables}
